---
title: "Preliminary Results"
author: "Tiff Stephens"
date: "12/1/2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
library(knitr); library(dplyr); library(tidyr); library(ggplot2); library(DT); library(cowplot); library(leaflet)
theme_set(theme_classic())
```


### 1. Import cleaned .csv file and look at dataframe.
```{r, include = FALSE}
# read clean file from github
cr.mat <- read.csv("https://raw.githubusercontent.com/APECS-ak/sediment_2017/master/CLEAN_core_matter.csv", stringsAsFactors = FALSE, header = TRUE)

# check 
str(cr.mat)
```


### 2. Remove excess columns. <br>

#### (e.g. pan + crucible data recorded prior to final calcualted weights)
```{r}

# remove pan, crucible, and sediment combo weights not necessary for analysis
cr.mat = cr.mat %>% 
  select(-na.rm, -pan_number, -pan_weight, -pan_sed_ww_pre_60C, -pan_sed_dw_post_60C, -pan_sed_dw_pre_450C, -pan_sed_dw_post450C, -crucible_weight, -crucible_sed_dw_pre_950C, -crucible_sed_dw_post_950C, -sed_ww_pre_60C, -sed_dw_post_60C, -sed_dw_pre_450C, -sed_dw_post_450C, -sed_dw_pre_950C, -sed_dw_post_950C)


colnames(cr.mat)
```

<br>

### 3. Where are the processed cores from?
```{r}
# new data frame for all cores
locations_all = cr.mat %>%
  distinct(cr.mat, latitude, longitude, site) %>%
  drop_na()

# map "locations_all"
leaflet(locations_all) %>%
  addTiles() %>%
  addMarkers(~longitude, ~latitude, popup = ~site, label = ~site)
```

<br>

### 4. Plotting
#### All sites
```{r, fig.height = 10, fig.width = 8, echo = FALSE, warning = FALSE}
levels=c("Low","Mid","High") # order zones

# plot organic matter
plot_om_all = ggplot(cr.mat, aes(layer_number, om_density)) +
  stat_summary(fun.y=mean, geom="point")+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.5) +
  coord_flip() +
  scale_x_reverse() +
  xlab("Core Depth") +
  ylab("Organic Matter (g cm^3)") +
  #scale_x_continuous(position = "top") +
  #scale_y_continuous(position = "left") +
  facet_wrap(~site, ncol = 1)

# plot inorganic matter
plot_im_all = ggplot(cr.mat, aes(layer_number, im_density)) +
  stat_summary(fun.y=mean, geom="point")+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.5) +
  coord_flip() +
  scale_x_reverse() +
  xlab("Core Depth") +
  ylab("Organic Matter (g cm^3)") +
  #scale_x_continuous(position = "top") +
  #scale_y_continuous(position = "left") +
  facet_wrap(~site, ncol = 1)

# organize om and im side-by-side
plot_grid(plot_om_all, plot_im_all, align = "hv", ncol = 2)
```

<br>

#### You will notice that site L06 is way different than the rest, so far. This core was loaded with chunky woody debris; it's real data but an outlier compared to the other sites. Here are two images of the sliced core from site L06:

```{r, include = FALSE}
img1_path = "images/woodchunk1.JPG"
img2_path = "images/woodchunk2.JPG"
```

```{r out.width = "50%", echo = FALSE}
include_graphics(img1_path)
include_graphics(img2_path)
```


<br>

#### For the sake of looking at regional patterns, let's remove site L06 until we have more sites processed:

```{r}
cr.mat_noL06 = cr.mat[-c(1:30),] # remove L06 (for now) because organic matter is off the scale
```

<br>
<br>


#### Here are the data, by site, without L06:
```{r, fig.height = 10, fig.width = 8, echo = FALSE, warning = FALSE}
levels=c("Low","Mid","High") # order zones

# plot organic matter
plot_om_all = ggplot(cr.mat_noL06, aes(layer_number, om_density)) +
  stat_summary(fun.y=mean, geom="point")+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.5) +
  coord_flip() +
  scale_x_reverse() +
  xlab("Core Depth") +
  ylab("Organic Matter (g cm^3)") +
  #scale_x_continuous(position = "top") +
  #scale_y_continuous(position = "left") +
  facet_wrap(~site, ncol = 1)

# plot inorganic matter
plot_im_all = ggplot(cr.mat_noL06, aes(layer_number, im_density)) +
  stat_summary(fun.y=mean, geom="point")+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.5) +
  coord_flip() +
  scale_x_reverse() +
  xlab("Core Depth") +
  ylab("Organic Matter (g cm^3)") +
  #scale_x_continuous(position = "top") +
  #scale_y_continuous(position = "left") +
  facet_wrap(~site, ncol = 1)

# organize om and im side-by-side
plot_grid(plot_om_all, plot_im_all, align = "hv", ncol = 2)
```








<br>
<br>

#### Here are the data grouped by sea otter zone (witout L06)
```{r fig2, fig.height = 10, fig.width = 8, echo = FALSE, warning = FALSE}

cr.mat_noL06$otter_region=factor(cr.mat_noL06$otter_region, levels=c("Low","Mid","High")) # order zones


# plot organic matter
plot_om = ggplot(cr.mat_noL06, aes(layer_number, om_density)) +
  stat_summary(fun.y=mean, geom="point")+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.5) +
  coord_flip() +
  scale_x_reverse() +
  xlab("Core Depth") +
  ylab("Organic Matter (g cm^3)") +
  #scale_x_continuous(position = "top") +
  #scale_y_continuous(position = "left") +
  facet_wrap(~otter_region, ncol = 1)

#plopt inorganic matter
plot_im = ggplot(cr.mat_noL06, aes(layer_number, im_density)) +
  stat_summary(fun.y=mean, geom="point")+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.5) +
  coord_flip() +
  scale_x_reverse() +
  xlab("Core Depth") +
  ylab("Inorganic Matter (g cm^3)") +
  #scale_x_continuous(position = "top") +
  #scale_y_continuous(position = "left") +
  facet_wrap(~otter_region, ncol = 1)

# organize om and im side-by-side
plot_grid(plot_om, plot_im, align = "hv", ncol = 2)
```


```{r, echo = FALSE, message = FALSE}
# make new dataframe for mapping
#locations = cr.mat_noL06 %>%
  #distinct(cr.mat_noL06, latitude, longitude, site) %>%
  #drop_na()

# turn into leaflet map
#leaflet(locations) %>%
  #addTiles() %>%
  #addMarkers(~longitude, ~latitude, popup = ~site, label = ~site)
```


